Latitudinal gradients in niche breadth vary across ecosystem types.

### 2015-09-01 manuscript/AC-reply-to-referees.tex

% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------
{\Large \bf Reply to Reviewer 2}
% ---------------------------------------

  The Reviewer stated that our study is interesting and covers ``a timely and
  important topic. The core analysis of the paper (scaling relationships of
  food web metrics with species richness) are convincingly presented and
  comprehensively discussed.'' Nevertheless, the Reviewer has suggested several
  ways in which our manuscript could be improved. We have responded to each suggestion
  in turn, and believe that our replies (preceded by \textbf{R}) have further improved
  the manuscript.

% ---------------------------------------

  1) Network terminology and methods might be inaccessible to researchers
  without network background.

@@

  \begin{refquote}

    Please re-label the y-axis on Fig. 1 to improve clarity. The top predators
    are missing in this figure although they are mentioned in the legend.

  \end{refquote}

  \textbf{R:} The top predators were indeed included in this figure (we suspect that the reviewer
  may have experienced a printing error), but as we have followed the reviewer's earlier
  suggestion and moved the by-trophic-level analysis to the supplemental information this becomes
  a moot point. We have slightly modified the y-axis on the new version of this figure and feel that
  it is now much clearer.

### 2015-09-02 manuscript/latitudepaper.tex
Tried to throw something together about freshwaters

\title{Specialisation in food webs scales with species richness but not with latitude}
\author{Alyssa R. Cirtwill$^{1,2}$, Daniel B. Stouffer$^{1}$, Tamara N. Romanuk$^{2}$}
\date{\small$^1$Centre for Integrative Ecology\\School of Biological Sciences\\University of Canterbury\\
Private Bag 4800\\Christchurch 8140, New Zealand \\
\medskip$^2$Department of Biology\\
Life Science Centre, Dalhousie University\\1355 Oxford St., P0 BOX 15000\\
Halifax NS, B3H 4R2, Canada\\}

@@

  The tendency of food-web structure to exhibit scaling relationships with
  species richness has been well-established~\cite{Dunne2004,Riede2010}. As
  species richness in particular is also known to vary systematically over
  latitude~\cite{Kaufman1995,Macpherson2002,Hillebrand2004,Schemske2009},
  intuitively one might suspect that any relationship between food-web
  properties such as generality might be due to the latitudinal gradient in
  species richness. In this dataset, however, we did not find overall
  latitudinal gradients in species richness, links per species,generality,
  vulnerability, or the proportions of food webs accounted for by basal
  resources, intermediate consumers, and top predators, except in lake and
  stream food webs where the proportions of top predators tended to decrease
  towards the poles  (see \emph{Appendix S3} for details).

  The lack of a latitudinal gradient in species richness in this dataset
  contrasts strongly with other
  studies~\cite{Kaufman1995,Macpherson2002,Hillebrand2004,Schemske2009}. As
  numbers of species and links included in a food web vary strongly with
  sampling effort as well as with the underlying diversity of the study area,
  it is possible that the lack of latitudinal trends here is a result of
  researchers tending to expend similar amounts of sampling effort across
  studies. This could result in food webs describing species-rich tropical
  communities omitting more species and links than species-poor arctic
  communities. In addition, it is worth noting that gradients in species
  richness are generally measured for a single taxonomic group at a
  time~\cite{Kaufman1995,Macpherson2002,Hillebrand2004,Schemske2009}. It is possible
  that these taxa are not well-represented in our food webs and that the
  dominant taxa in them do not have an underlying latitudinal gradient in
  richness.In either case, the lack of a strong association between species
  richness and latitude in any ecosystem type means that any effect of
  latitude on other scaling relationships is not being driven by an underlying
  latitudinal gradient in species richness. The lack of confounding effects of
  latitude allows us to more clearly assess effects of latitude on scaling
  with species richness and proportions of species in different trophic
  levels.

@@

  Unlike other ecosystem types, the scaling of generality in lake and stream
  food webs was stronger (i.e.,generality increased more steeply with
  increasing species richness) in higher-latitude food webs. In lake food webs,
  this trend was echoed in the scaling relationships between species richness
  and links per species and vulnerability. This means that species in tropical
  freshwater communities gain fewer additional feeding links per additional
  species in the web and that species in tropical lakes also gain fewer
  predators, and fewer links in general, per additional species than species
  in high-latitude lakes. These trends are consistent with the hypothesis that
  greater stability in the tropics leads to narrower niches~\cite{Vazquez2004}
  and a higher proportion of specialists. This may be partly due to high-latitude 
  species tending to switch between different seasonally-available 
  prey~\cite{Magalhaes1993,Wilhelm1999,Isaac2012} while tropical
  freshwater ecosystems may have more stable composition.

  The scaling of link density, generality and vulnerability with proportions
  of species at different trophic levels, however, did not differ between
  lakes and most other food webs. Indeed, there was comparatively little
  variation in scaling with the trophic-level breakdown of a web across
  ecosystem types. Nevertheless, the negative relationships between the
  proportion of top predators in a web and link density and vulnerability were
  weaker (i.e., the scaling exponent was more strongly negative) in low-latitude 
  stream ecosystems. This may imply that top predators in high-latitude 
  streams tend to be more generalist than their low-latitude
  equivalents~\cite{Winemiller2008}, perhaps due to prey-switching during
  seasonal food shortages~\cite{Magalhaes1993}. More generally, the lack of
  correspondence between the scaling of food-web properties with species
  richness and with proportions of species at each trophic level suggests that
  the size and trophic breakdowns of a community can each provide different
  information~\cite{Downing2002}.

@@

  We thank members of the Romanuk, Stouffer, and Tylianakis labs for their comments on
  the manuscript.

\section*{Funding}
  This research was supported by an NSERC USRA undergraduate scholarship and NSERC PGS-D 
  graduate scholarship (to ARC), as well as by a Rutherford XXXYYY (to DBS).

\newpage
\bibliographystyle{vancouver}%Compile with vancouver.bst style file
\bibliography{noISN}% your .bib file(s)

>>>>

\title{Latitudinal gradients in niche breadth vary across ecosystem types.}
\author{Alyssa R. Cirtwill$^{1,2}$, Daniel B. Stouffer$^{1}$, Tamara N. Romanuk$^{2}$}
\date{\small$^1$Centre for Integrative Ecology\\School of Biological Sciences\\University of Canterbury\\
Private Bag 4800\\Christchurch 8140, New Zealand \\
\medskip$^2$Department of Biology\\
Life Science Centre, Dalhousie University\\1355 Oxford St., P0 BOX 15000\\
Halifax NS, B3H 4R2, Canada\\}

@@

  The tendency of food-web structure to exhibit scaling relationships with
  species richness has been well-established~\cite{Dunne2004,Riede2010}. As
  species richness in particular is also known to vary systematically over
  latitude~\cite{Kaufman1995,Macpherson2002,Hillebrand2004,Schemske2009},
  intuitively one might suspect that any relationship between food-web
  properties such as generality might be due to the latitudinal gradient in
  species richness. In this dataset, however, we did not find overall
  latitudinal gradients in species richness, links per species, generality, or
  vulnerability.
  % , or the proportions of food webs accounted for by basal
  % resources, intermediate consumers, and top predators, except in lake and
  % stream food webs where the proportions of top predators tended to decrease
  % towards the poles  (see \emph{Appendix S3} for details).

  The lack of a latitudinal gradient in species richness in this dataset
  contrasts strongly with other
  studies~\cite{Kaufman1995,Macpherson2002,Hillebrand2004,Schemske2009}. As
  numbers of species and links included in a food web vary strongly with
  sampling effort as well as with the underlying diversity of the study area,
  it is possible that the lack of latitudinal trends here is a result of
  researchers tending to expend similar amounts of sampling effort across
  studies. This could result in food webs describing species-rich tropical
  communities omitting more species and links than studies of species-poor arctic
  communities. In addition, it is worth noting that gradients in species
  richness are generally measured for a single taxonomic group at a
  time~\cite{Kaufman1995,Macpherson2002,Hillebrand2004,Schemske2009}. It is possible
  that these taxa are not well-represented in our food webs and that the
  dominant taxa in them do not have an underlying latitudinal gradient in
  richness. In either case, the lack of a strong association between species
  richness and latitude in any ecosystem type means that any effect of
  latitude on other scaling relationships is not being driven by an underlying
  latitudinal gradient in species richness. The lack of confounding effects of
  latitude allows us to more clearly assess effects of latitude on scaling
  with species richness and proportions of species in different trophic
  levels.

@@

  Unlike other ecosystem types, the scaling of generality in lake and stream
  food webs was stronger (i.e.,generality increased more steeply with
  increasing species richness) in higher-latitude food webs. In lake food webs,
  this trend was echoed in the scaling relationships between species richness
  and links per species and vulnerability. This means that species in tropical
  freshwater communities gain fewer additional feeding links per additional
  species in the web and that species in tropical lakes also gain fewer
  predators, and fewer links in general, per additional species than species
  in high-latitude lakes. These trends are consistent with the hypothesis that
  greater stability in the tropics leads to narrower niches~\cite{Vazquez2004}
  and a higher proportion of specialists. That freshwater food webs supported
  the hypothesis of narrower niches in the tropics, while other ecosystem types
  did not, is particularly exciting given that these ecosystems (especially streams)
  are known for being highly variable and that seasonal variability is one of the
  proposed drivers of the latitude-niche breadth hypothesis~\cite{Vazquez2004}.

  Both streams and lakes can experience severe changes in water temperature and volume
  (e.g., floods, drying, freezing) that remove food or other resources (notably oxygen
  during freezing events~\cite{})~\cite{Winterbourn1997, one for lakes}. These events
  are often linked to seasonal events such as snowmelts (although not always-- New Zealand
  streams are a notable exception~\cite{Winterbourn1997}). Further, both temperate streams 
  and lakes tend to experience seasonal strong pulses of allochthonous inputs (e.g., fallen leaves, 
  terrestrial invertebrates~\cite{Nakano2001,Zeng2008,}; again New Zealand streams are an
  exception to this trend~\cite{Winterbourn1997}). These trends combined mean that,
  relative to larger, more abiotically stable estuarine and marine communities, freshwater
  food webs may experience high turnover in both community composition and productivity.
  Even in New Zealand streams, which experience unpredictable flooding and do not typically
  receive seasonal allochthonous subsidies, more frequent disturbance is associated with lower
  diversity and density of insect larvae, and lower species richness~\cite{Winterbourn1997}.

  High variation in the abundances of different resources could select for more generalist consumers
  in temperature freshwater communities, as seen in our dataset. In this case the non-seasonality of
  New Zealand streams, which account for 31/71 of the stream webs in our dataset, supports this
  hypothesis as the New Zealand stream fauna are known to be ecological generalists,
  probably as a result of variability of their habitat~\cite{Winterbourn1997}. In other freshwater
  communities the sequence of available resources might be more predictable, but the ability to 
  switch between different prey is not likely to be less important. Importantly, while terrestrial
  communities are also strongly seasonal at high latitudes, and can also obtain significant allochthonous
  inputs~\cite{Nakano2001}, terrestrial consumers tend to be more morphologically specialized for
  feeding on particular prey~\cite{Liem1990}. This means that primarily gape-limited aquatic consumers
  tend to be more general across all types of aquatic habitats~\cite{Liem1990,Shurin2006}.
  The key to this this explanation of the differences between freshwater and marine and estuarine 
  ecosystems is whether the former experience more severe seasonal variation. Although we are not
  aware of any study explicitly comparing seasonal variation in freshwater and saltwater or brackish
  food webs in a similar location, we believe that freshwater ecosystems are indeed likely to 
  experience more severe changes because of their small size. While oceans and estuaries certainly
  do vary in terms of water temperature and nutrients over the course of a year~\cite{}, because they
  are not isolated in the midst of a terrestrial matrix these changes are buffered. Further, estuaries
  and especially marine environments do not typically freeze over or dry up completely-- these dramatic
  changes are unique to freshwater communities.

  % This may be partly due to high-latitude 
  % species tending to switch between different seasonally-available 
  % prey~\cite{Magalhaes1993,Wilhelm1999,Isaac2012} while tropical
  % freshwater ecosystems may have more stable composition.

  % The scaling of link density, generality and vulnerability with proportions
  % of species at different trophic levels, however, did not differ between
  % lakes and most other food webs. Indeed, there was comparatively little
  % variation in scaling with the trophic-level breakdown of a web across
  % ecosystem types. Nevertheless, the negative relationships between the
  % proportion of top predators in a web and link density and vulnerability were
  % weaker (i.e., the scaling exponent was more strongly negative) in low-latitude 
  % stream ecosystems. This may imply that top predators in high-latitude 
  % streams tend to be more generalist than their low-latitude
  % equivalents~\cite{Winemiller2008}, perhaps due to prey-switching during
  % seasonal food shortages~\cite{Magalhaes1993}. More generally, the lack of
  % correspondence between the scaling of food-web properties with species
  % richness and with proportions of species at each trophic level suggests that
  % the size and trophic breakdowns of a community can each provide different
  % information~\cite{Downing2002}.

@@

  We thank members of the Romanuk, Stouffer, and Tylianakis labs for their comments on
  the manuscript. We also thank Angus McIntosh for valuable discussion on stream food webs.

\section*{Funding}
  This research was supported by an NSERC USRA undergraduate scholarship and NSERC PGS-D 
  graduate scholarship (to ARC), a Rutherford XXXYYY (to DBS), and an NSERC Discovery Grant to TNR.

1
\newpage
\bibliographystyle{vancouver}%Compile with vancouver.bst style file
\bibliography{noISN}% your .bib file(s)
